---
title: "Cirsium head-orientation macroecology"
subtitle: "Global environmental analysis and Japan pollinator–environment analysis"
author: "ZHANG RACHEL"
date: "`r format(Sys.Date(), '%Y-%m-%d')`"
output:
  html_document:
    toc: true
    toc_depth: 3
    number_sections: true
    theme: flatly
    df_print: paged
params:
  root_dir: "C:/Users/zuizui/cirsium_inat"
  run_global_analysis: false
  run_japan_analysis: false
  show_figures: true
  max_table_rows: 30
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.width = 8,
  fig.height = 5,
  dpi = 300
)

pkgs <- c(
  "dplyr", "readr", "stringr", "tibble", "tidyr",
  "ggplot2", "forcats", "knitr", "kableExtra", "patchwork",
  "scales"
)

for (p in pkgs) {
  if (!requireNamespace(p, quietly = TRUE)) {
    install.packages(p, dependencies = TRUE)
  }
}

suppressPackageStartupMessages({
  library(dplyr)
  library(readr)
  library(stringr)
  library(tibble)
  library(tidyr)
  library(ggplot2)
  library(forcats)
  library(knitr)
  library(kableExtra)
  library(scales)
})

theme_report <- function(base_size = 11) {
  ggplot2::theme_classic(base_size = base_size) +
    ggplot2::theme(
      axis.line = element_line(linewidth = 0.35),
      axis.ticks = element_line(linewidth = 0.35),
      panel.grid.major.x = element_line(colour = "grey88", linewidth = 0.25),
      panel.grid.major.y = element_blank(),
      strip.background = element_rect(fill = "grey92", colour = "grey60", linewidth = 0.35),
      strip.text = element_text(face = "bold"),
      legend.title = element_text(face = "bold")
    )
}

safe_read_csv <- function(path, required = FALSE) {
  if (!file.exists(path)) {
    if (required) stop("Missing file: ", path)
    return(NULL)
  }
  readr::read_csv(path, show_col_types = FALSE)
}

safe_kable <- function(x, caption = NULL, n = params$max_table_rows) {
  if (is.null(x) || nrow(x) == 0) {
    cat("*No table found.*\n\n")
    return(invisible(NULL))
  }

  x |>
    head(n) |>
    knitr::kable(caption = caption, digits = 4) |>
    kableExtra::kable_styling(full_width = FALSE)
}

find_first_file <- function(root, pattern, recursive = TRUE) {
  if (!dir.exists(root)) return(NA_character_)
  hits <- list.files(root, pattern = pattern, recursive = recursive, full.names = TRUE, ignore.case = TRUE)
  if (length(hits) == 0) NA_character_ else hits[1]
}

include_if_exists <- function(path, caption = NULL, width = "100%") {
  if (!isTRUE(params$show_figures)) return(invisible(NULL))
  if (is.na(path) || !file.exists(path)) {
    cat("*Figure not found:* `", path, "`\n\n", sep = "")
    return(invisible(NULL))
  }
  if (!is.null(caption)) cat("**", caption, "**\n\n", sep = "")
  knitr::include_graphics(path)
}

pretty_var <- function(x) {
  x0 <- stringr::str_remove(as.character(x), "^z_")
  dplyr::case_when(
    x0 == "lon" ~ "Longitude",
    x0 == "lat" ~ "Latitude",
    x0 == "z_lon" ~ "Longitude",
    x0 == "z_lat" ~ "Latitude",
    x0 == "z_lon2" ~ "Longitude²",
    x0 == "z_lat2" ~ "Latitude²",
    x0 == "z_lon_lat" ~ "Longitude × latitude",
    x0 == "longitude" ~ "Longitude",
    x0 == "latitude" ~ "Latitude",
    x0 == "elevation" ~ "Elevation",
    x0 == "slope" ~ "Slope",
    x0 == "roughness" ~ "Roughness",
    x0 == "soil_bdod_0_30cm" ~ "Soil bulk density",
    x0 == "soil_cec_0_30cm" ~ "Soil CEC",
    x0 == "soil_cfvo_0_30cm" ~ "Soil coarse fragments",
    x0 == "soil_clay_0_30cm" ~ "Soil clay",
    x0 == "soil_sand_0_30cm" ~ "Soil sand",
    x0 == "soil_silt_0_30cm" ~ "Soil silt",
    x0 == "soil_nitrogen_0_30cm" ~ "Soil nitrogen",
    x0 == "soil_phh2o_0_30cm" ~ "Soil pH",
    x0 == "soil_soc_0_30cm" ~ "Soil organic carbon",
    x0 == "soil_ocd_0_30cm" ~ "Soil organic carbon density",
    TRUE ~ x0 |>
      stringr::str_replace_all("_", " ") |>
      stringr::str_replace_all("BIO", "BIO")
  )
}

root_dir <- params$root_dir
```

# Report purpose

This R Markdown document is a **clean report wrapper** for the two final R analyses:

1. **Global analysis**  
   Machine-learning-predicted Cirsium head orientation + CHELSA climate + topography + SoilGrids.  
   The final R script runs RF, correlation clustering, VIF pruning, common-N weighted GLM, LRT, GAM, and publication-level figures.

2. **Japan analysis**  
   Japan-scale Cirsium trait grid + environmental PCA + pollinator SDMs.  
   The final R script compares pollinator grouping schemes and fits binomial / beta-binomial models for the grid-level proportion of nodding-flowered species.

The goal is not to duplicate every line of the long scripts, but to:

- run the final scripts when requested,
- read their main outputs,
- summarize model comparisons and key coefficients,
- display publication-ready figures,
- keep the analysis order clear.

# Main scripts

```{r paths}
global_script <- file.path(root_dir, "global_cirsium_full_rerun_RF_GLM_GAM_publication_figures.R")
japan_script  <- file.path(root_dir, "cirsium_pollinator_grouping_sensitivity_model_comparison_full_fixed.R")

global_out_dir <- file.path(root_dir, "global_cirsium_FULL_RERUN_RF_GLM_GAM_PUBLICATION")
global_rf_dir <- file.path(global_out_dir, "01_RF")
global_model_dir <- file.path(global_out_dir, "02_GLM_GAM")
global_pub_dir <- file.path(global_out_dir, "04_publication_figures")
global_pub_png <- file.path(global_pub_dir, "png")
global_pub_data <- file.path(global_pub_dir, "figure_data")

japan_root_dir <- file.path(root_dir, "chelsa_pollinator_enmeval_rebuild_no_swe")
japan_out_dir <- file.path(japan_root_dir, "cirsium_pollinator_grouping_sensitivity_model_comparison")
japan_fig_dir <- file.path(japan_out_dir, "figures")
japan_pub_dir <- file.path(japan_out_dir, "publication_figures_best_model")

tibble::tibble(
  item = c(
    "Global script",
    "Japan script",
    "Global output directory",
    "Japan output directory"
  ),
  path = c(global_script, japan_script, global_out_dir, japan_out_dir),
  exists = file.exists(path) | dir.exists(path)
) |>
  knitr::kable()
```

# Optional full rerun

By default, this report **does not rerun** the long analyses.  
To rerun them while knitting, set the YAML parameters:

```yaml
params:
  run_global_analysis: true
  run_japan_analysis: true
```

```{r run-global, eval=params$run_global_analysis}
stopifnot(file.exists(global_script))
source(global_script)
```

```{r run-japan, eval=params$run_japan_analysis}
stopifnot(file.exists(japan_script))
source(japan_script)
```

# Global analysis: head orientation vs environment

## Analysis workflow

The global script is organized as:

1. Load the enriched table with predicted head orientation and extracted environmental variables.
2. Run RF without spatial predictors.
3. Run RF with spatial predictors.
4. Select RF-top predictors.
5. Reduce multicollinearity using correlation clustering.
6. Apply VIF pruning.
7. Fit common-N weighted logistic GLMs.
8. Test incremental contributions of environment and space by likelihood-ratio tests.
9. Fit GAMs.
10. Export publication-level figures in PNG/PDF/TIFF.

The response variable is binary:

- `upward = 0`
- `nodding = 1`

Class imbalance is handled by:

- balanced resampling for RF,
- class-balanced `case_weight` for GLM/GAM.

## Input and class balance

```{r global-input}
global_enriched <- file.path(
  root_dir,
  "global_cirsium_CHELSA_SOILGRIDS_TOPO_RF_GLM_GAM",
  "01_orientation_points_CHELSA_topography_soilgrids.csv"
)

global_dat <- safe_read_csv(global_enriched)

if (!is.null(global_dat)) {
  if ("orientation_factor" %in% names(global_dat)) {
    global_dat |>
      count(orientation_factor) |>
      mutate(prop = n / sum(n)) |>
      safe_kable("Global orientation records")
  } else if ("y_nodding" %in% names(global_dat)) {
    global_dat |>
      mutate(orientation_factor = ifelse(y_nodding == 1, "nodding", "upward")) |>
      count(orientation_factor) |>
      mutate(prop = n / sum(n)) |>
      safe_kable("Global orientation records")
  }
} else {
  cat("*Global enriched input table not found. Run the global script first or check the path.*\n")
}
```

## RF variable importance

```{r global-rf}
rf_imp <- safe_read_csv(
  file.path(global_rf_dir, "occurrence_CHELSA_TOPO_SOIL_RF_importance_repeats_summary.csv")
)

safe_kable(rf_imp, "Top RF importance without space")

if (!is.null(rf_imp)) {
  rf_imp |>
    slice_max(mean_MDA, n = 25, with_ties = FALSE) |>
    mutate(variable_label = fct_reorder(pretty_var(variable), mean_MDA)) |>
    ggplot(aes(x = mean_MDA, y = variable_label)) +
    geom_col(width = 0.72) +
    geom_errorbar(
      aes(xmin = mean_MDA - sd_MDA, xmax = mean_MDA + sd_MDA),
      orientation = "y",
      width = 0.15,
      linewidth = 0.35
    ) +
    theme_report(10) +
    labs(
      x = "Random forest importance\n(mean decrease accuracy)",
      y = NULL
    )
}
```

## Final predictors after correlation clustering and VIF pruning

```{r global-vif}
final_vars <- safe_read_csv(file.path(global_model_dir, "04_final_variables_for_GLM_GAM.csv"))
vif_tbl <- safe_read_csv(file.path(global_model_dir, "03_VIF_final_table.csv"))
vif_removed <- safe_read_csv(file.path(global_model_dir, "03_VIF_removed_variables.csv"))

safe_kable(final_vars, "Final variables used in GLM/GAM")
safe_kable(vif_tbl, "Final VIF table after pruning")
safe_kable(vif_removed, "Variables removed by VIF pruning")
```

## GLM model comparison

```{r global-glm-comparison}
glm_comp <- safe_read_csv(file.path(global_model_dir, "05_GLM_model_comparison.csv"))

safe_kable(glm_comp, "Global GLM model comparison")

if (!is.null(glm_comp)) {
  glm_comp |>
    arrange(AIC) |>
    mutate(
      model_label = case_when(
        model == "env_topography_soil_new_RFtop_VIF" ~ "Environment + topography + soil",
        model == "space_only_commonN" ~ "Space only",
        model == "env_topography_soil_plus_space_new_RFtop_VIF" ~ "Environment + topography + soil + space",
        TRUE ~ model
      ),
      model_label = fct_reorder(model_label, delta_AIC)
    ) |>
    ggplot(aes(x = delta_AIC, y = model_label)) +
    geom_col(width = 0.72) +
    theme_report(11) +
    labs(x = expression(Delta*AIC), y = NULL)
}
```

## GLM coefficients

```{r global-glm-coefficients}
glm_coef <- safe_read_csv(file.path(global_model_dir, "05_GLM_coefficients.csv"))

safe_kable(glm_coef, "Global GLM coefficients")

if (!is.null(glm_coef)) {
  coef_plot <- glm_coef |>
    filter(model == "env_topography_soil_plus_space_new_RFtop_VIF", term != "(Intercept)") |>
    mutate(
      conf.low = ifelse("conf.low" %in% names(glm_coef), conf.low, estimate - 1.96 * std.error),
      conf.high = ifelse("conf.high" %in% names(glm_coef), conf.high, estimate + 1.96 * std.error),
      term_label = fct_reorder(pretty_var(term), estimate)
    )

  coef_plot |>
    ggplot(aes(x = estimate, y = term_label)) +
    geom_vline(xintercept = 0, linetype = 2, linewidth = 0.35) +
    geom_errorbar(
      aes(xmin = conf.low, xmax = conf.high),
      orientation = "y",
      width = 0.15,
      linewidth = 0.42
    ) +
    geom_point(size = 2.2) +
    theme_report(10.5) +
    labs(x = "Weighted logistic coefficient", y = NULL)
}
```

## Incremental GLM tests

```{r global-lrt}
glm_lrt <- safe_read_csv(file.path(global_model_dir, "06_GLM_incremental_tests.csv"))
safe_kable(glm_lrt, "Incremental GLM likelihood-ratio tests")
```

## GAM comparison and smooth terms

```{r global-gam}
gam_comp <- safe_read_csv(file.path(global_model_dir, "07_GAM_model_comparison.csv"))
gam_smooth <- safe_read_csv(file.path(global_model_dir, "07_GAM_smooth_terms.csv"))

safe_kable(gam_comp, "Global GAM model comparison")
safe_kable(gam_smooth, "Global GAM smooth terms")

if (!is.null(gam_comp)) {
  gam_comp |>
    mutate(
      model_label = case_when(
        model == "GAM_env_topography_soil" ~ "GAM: environment",
        model == "GAM_env_topography_soil_plus_space" ~ "GAM: environment + space",
        TRUE ~ model
      ),
      model_label = fct_reorder(model_label, delta_AIC)
    ) |>
    ggplot(aes(x = delta_AIC, y = model_label)) +
    geom_col(width = 0.72) +
    theme_report(11) +
    labs(x = expression(Delta*AIC), y = NULL)
}
```

## Global publication figures

```{r global-figures, results='asis'}
global_fig_index <- safe_read_csv(file.path(global_pub_dir, "00_publication_figure_index.csv"))
safe_kable(global_fig_index, "Global publication figure index")

main_fig1 <- file.path(global_pub_png, "MainFig1_RF_GLM_model_selection_coefficients.png")
main_fig2 <- file.path(global_pub_png, "MainFig2_response_curves_prediction_map.png")
main_fig3 <- file.path(global_pub_png, "MainFig3_GAM_summary.png")

include_if_exists(main_fig1, "MainFig1: RF, GLM model selection, and coefficients")
include_if_exists(main_fig2, "MainFig2: GLM response curves and prediction map")
include_if_exists(main_fig3, "MainFig3: GAM summary")
```

# Japan analysis: Cirsium grid proportion vs environment and pollinators

## Analysis workflow

The Japan script compares how different pollinator grouping schemes affect models of the **regional proportion of nodding-flowered Cirsium species**.

Response:

```r
cbind(n_nodding_species, n_upward_species)
```

Main predictor blocks:

- grid-level environmental PCA from candidate environmental rasters,
- pollinator SDM groups,
- spatial structure.

Pollinator grouping schemes:

1. `broad4_direct`: bee + butterfly + fly + hawkmoth
2. `lepidoptera3_direct`: bee + Lepidoptera + fly
3. `order3_direct`: Hymenoptera + Lepidoptera + Diptera
4. `order3_pca`: PCA of Hymenoptera / Lepidoptera / Diptera
5. `bombus_vs_nonbombus_all`: Bombus + all non-Bombus pollinators
6. `bombus_vs_nonbombus_bees`: Bombus + non-Bombus bees only
7. `total_pollinator`: all pollinator SDMs merged

Main model families:

- binomial GLM,
- beta-binomial GLMM via `glmmTMB`.

## Input and grid response

```{r japan-input}
japan_grid_candidates <- c(
  file.path(japan_out_dir, "scheme_order3_pca", "analysis_grid_data.csv"),
  file.path(japan_out_dir, "scheme_order3_direct", "analysis_grid_data.csv"),
  file.path(japan_out_dir, "scheme_broad4_direct", "analysis_grid_data.csv")
)

japan_grid_file <- japan_grid_candidates[file.exists(japan_grid_candidates)][1]
japan_grid <- safe_read_csv(japan_grid_file)

if (!is.null(japan_grid)) {
  japan_grid |>
    summarise(
      n_grids = n(),
      mean_total_species = mean(n_total_species, na.rm = TRUE),
      mean_prop_nodding = mean(prop_nodding_species, na.rm = TRUE),
      min_prop_nodding = min(prop_nodding_species, na.rm = TRUE),
      max_prop_nodding = max(prop_nodding_species, na.rm = TRUE)
    ) |>
    safe_kable("Japan grid-level response summary")
} else {
  cat("*Japan grid data not found. Run the Japan script first or check the path.*\n")
}
```

## Model comparison across pollinator grouping schemes

```{r japan-model-comparison}
bb_comp <- safe_read_csv(file.path(japan_out_dir, "all_scheme_beta_binomial_model_comparison.csv"))
glm_comp_jp <- safe_read_csv(file.path(japan_out_dir, "all_scheme_binomial_GLM_model_comparison.csv"))
best_models <- safe_read_csv(file.path(japan_out_dir, "all_scheme_best_models.csv"))

safe_kable(bb_comp, "Japan beta-binomial model comparison")
safe_kable(glm_comp_jp, "Japan binomial GLM model comparison")
safe_kable(best_models, "Best models by scheme")

if (!is.null(bb_comp)) {
  bb_plot <- bb_comp |>
    filter(is.finite(AIC)) |>
    arrange(AIC) |>
    mutate(
      global_delta_AIC = ifelse(
        "global_delta_AIC" %in% names(bb_comp),
        global_delta_AIC,
        AIC - min(AIC, na.rm = TRUE)
      ),
      label = paste(scheme, model, sep = " / "),
      label = fct_reorder(label, global_delta_AIC)
    ) |>
    slice_head(n = 30)

  bb_plot |>
    ggplot(aes(x = global_delta_AIC, y = label)) +
    geom_col(width = 0.72) +
    theme_report(9) +
    labs(x = expression(Global~Delta*AIC), y = NULL)
}
```

## Best beta-binomial model

```{r japan-best}
if (!is.null(bb_comp)) {
  best_bb <- bb_comp |>
    filter(method == "beta_binomial_glmmTMB", is.finite(AIC)) |>
    arrange(AIC) |>
    slice(1)

  safe_kable(best_bb, "Global best beta-binomial model")

  best_scheme <- best_bb$scheme[1]
  best_model <- best_bb$model[1]
} else {
  best_scheme <- NA_character_
  best_model <- NA_character_
}

best_scheme
best_model
```

## Coefficients across schemes

```{r japan-coefficients}
coef_candidates <- c(
  file.path(japan_out_dir, "all_scheme_beta_binomial_coefficients.csv"),
  file.path(japan_out_dir, "all_scheme_coefficients.csv")
)

coef_file <- coef_candidates[file.exists(coef_candidates)][1]
jp_coef <- safe_read_csv(coef_file)

safe_kable(jp_coef, "Japan coefficients")

if (!is.null(jp_coef) && !is.na(best_scheme) && !is.na(best_model)) {
  jp_coef_plot <- jp_coef |>
    filter(
      scheme == best_scheme,
      model == best_model,
      method == "beta_binomial_glmmTMB",
      term != "(Intercept)",
      is.finite(estimate),
      is.finite(std.error)
    ) |>
    mutate(
      conf.low = estimate - 1.96 * std.error,
      conf.high = estimate + 1.96 * std.error,
      term_label = fct_reorder(term, estimate)
    )

  jp_coef_plot |>
    ggplot(aes(x = estimate, y = term_label)) +
    geom_vline(xintercept = 0, linetype = 2, linewidth = 0.35) +
    geom_errorbar(
      aes(xmin = conf.low, xmax = conf.high),
      orientation = "y",
      width = 0.15,
      linewidth = 0.42
    ) +
    geom_point(size = 2.2) +
    theme_report(10.5) +
    labs(x = "Beta-binomial coefficient", y = NULL)
}
```

## Predictor correlations

```{r japan-correlations}
corr_tbl <- safe_read_csv(file.path(japan_out_dir, "all_scheme_predictor_correlations.csv"))

safe_kable(corr_tbl, "Predictor correlations across schemes")

if (!is.null(corr_tbl)) {
  corr_tbl |>
    filter(abs_correlation >= 0.7) |>
    arrange(desc(abs_correlation)) |>
    mutate(pair = paste(var1, var2, sep = " — ")) |>
    slice_head(n = 30) |>
    mutate(pair = fct_reorder(pair, abs_correlation)) |>
    ggplot(aes(x = abs_correlation, y = pair)) +
    geom_col(width = 0.72) +
    geom_vline(xintercept = 0.7, linetype = 2) +
    geom_vline(xintercept = 0.85, linetype = 3) +
    theme_report(8.5) +
    labs(x = "|correlation|", y = NULL)
}
```

## PCA diagnostics

```{r japan-pca}
env_loadings <- safe_read_csv(file.path(japan_out_dir, "01_grid_preVIF_env_PCA_loadings.csv"))

if (is.null(env_loadings)) {
  env_loadings <- safe_read_csv(find_first_file(japan_out_dir, "env.*PCA.*loading|grid.*PCA.*loading"))
}

safe_kable(env_loadings, "Environmental PCA loadings")

if (!is.null(env_loadings)) {
  if (!"variable" %in% names(env_loadings)) {
    var_col <- intersect(c("var", "layer", "name"), names(env_loadings))[1]
    if (!is.na(var_col)) names(env_loadings)[names(env_loadings) == var_col] <- "variable"
  }

  if (any(stringr::str_detect(names(env_loadings), "^PC[0-9]+$"))) {
    env_long <- env_loadings |>
      select(variable, matches("^PC[0-9]+$")) |>
      pivot_longer(matches("^PC[0-9]+$"), names_to = "PC", values_to = "loading")
  } else if (all(c("PC", "loading") %in% names(env_loadings))) {
    env_long <- env_loadings |>
      select(variable, PC, loading)
  } else {
    env_long <- NULL
  }

  if (!is.null(env_long)) {
    env_top <- env_long |>
      mutate(abs_loading = abs(as.numeric(loading))) |>
      filter(PC %in% paste0("PC", 1:3)) |>
      group_by(PC) |>
      slice_max(abs_loading, n = 12, with_ties = FALSE) |>
      ungroup() |>
      mutate(variable = fct_reorder(variable, loading))

    env_top |>
      ggplot(aes(x = loading, y = variable)) +
      geom_vline(xintercept = 0, linetype = 2) +
      geom_col(width = 0.72) +
      facet_wrap(~ PC, scales = "free_y") +
      theme_report(9) +
      labs(x = "PCA loading", y = NULL)
  }
}
```

## Japan figures

```{r japan-figures, results='asis'}
japan_fig_index <- safe_read_csv(file.path(japan_pub_dir, "00_publication_figure_index.csv"))
safe_kable(japan_fig_index, "Japan publication figure index")

fig_candidates <- c(
  file.path(japan_pub_dir, "Fig1_global_model_comparison_top20.png"),
  file.path(japan_pub_dir, "Fig2_best_model_each_scheme.png"),
  file.path(japan_pub_dir, "Fig3_best_model_coefficients.png"),
  file.path(japan_pub_dir, "Fig4_pollinator_order_PCA_loadings.png"),
  file.path(japan_pub_dir, "Fig5_environmental_PCA_top_loadings.png"),
  file.path(japan_pub_dir, "Fig6_combined_prediction_residual_panel.png")
)

for (ff in fig_candidates) {
  include_if_exists(ff, caption = basename(ff))
}
```

# Integrated interpretation

## What the global analysis contributes

The global analysis tests whether ML-predicted nodding head orientation is associated with global-scale climate, topography, soil, and spatial structure.

Important interpretation points:

- RF is used for exploratory variable screening.
- Correlation clustering and VIF pruning reduce predictor collinearity.
- The final GLM is weighted to reduce class imbalance between upward and nodding records.
- Spatial polynomial terms help separate environmental association from broad geographic structure.
- GAMs provide a nonlinear check of the GLM patterns.

## What the Japan analysis contributes

The Japan analysis asks whether the **regional proportion of nodding-flowered species** is associated with:

- environmental gradients,
- pollinator SDM composition,
- spatial structure.

Important interpretation points:

- The response is grid-level species composition, not individual image records.
- Beta-binomial models are important because they can account for overdispersion in proportions.
- Pollinator grouping sensitivity analysis checks whether conclusions depend on how pollinators are grouped.
- Environmental PCA is used because grid-level candidate environmental predictors are highly correlated.

## How the two analyses connect

Together, the two analyses form a two-scale argument:

1. **Global analysis**  
   Nodding head orientation is associated with broad environmental gradients such as moisture, elevation, climate seasonality, and geography.

2. **Japan analysis**  
   Within Japan, the regional proportion of nodding-flowered Cirsium species is further examined against environmental gradients and pollinator community structure.

3. **Next step**  
   Phylogenetic mapping is needed to separate:
   - repeated environmental adaptation,
   - phylogenetic inertia,
   - historical biogeography,
   - unresolved local ecological mechanisms.

# Suggested next analyses

- Map head orientation onto a phylogeny.
- Test whether nodding orientation is clustered in specific clades.
- Estimate ancestral states and number of transitions.
- Compare environmental niches between upward and nodding lineages after phylogenetic correction.
- Validate functional hypotheses using pollinator observation and orientation manipulation experiments.
